\documentclass[11pt]{article}
\input{goalos_style.tex}
\input{goalos_figures.tex}

\begin{document}

\CanonicalCover
{GoalOS: Proof-Carrying Autonomy}
{The Canonical Proof-to-Capability Operating System for Autonomous AI Work}
{Institutional Research Monograph}
{GoalOS converts an objective into a bounded Proof Mission; unsupported claims into Proof Debt; Proof Debt into custom AGI Jobs; jobs into Evidence Dockets; evidence into Chronicle decisions; Chronicle-admitted results into a Validated Skill Graph; and admitted graph state into tamper-evident Merkle commitments. Future missions improve only from scoped, replayable, validated, root-verifiable capability.}

\begin{abstract}
Frontier models make candidate work abundant, but institutions require a governed method for deciding which work deserves trust, reuse, payment, memory, propagation, or rollback authority. This paper presents GoalOS as a proof-to-capability operating system for autonomous AI work. The canonical loop is: Objective $\rightarrow$ Proof Mission $\rightarrow$ AGI Jobs $\rightarrow$ Evidence $\rightarrow$ Chronicle $\rightarrow$ Validated Skill $\rightarrow$ On-chain Graph Root $\rightarrow$ Future Mission Improvement. Unsupported claims become explicit Proof Debt. An empty-by-design job factory converts that debt into mission-specific proof contracts. Returned ProofBundles populate an Evidence Docket. Chronicle acts as a memory firewall, admitting only capabilities that satisfy proof-level-specific evidence, replay, validation, scope, risk, utility, freshness, and rollback criteria. Admitted capabilities become nodes in a Validated Skill Graph rather than entries in a prompt library. A Merkle-native commitment layer binds typed roots for skills, evidence, ProofBundles, policies, attestations, encrypted capsules, and Chronicle decisions into append-auditable graph roots. Encryption, selective disclosure, zero-knowledge policy proofs, direct-agent authorization, and an explicit local/live boundary protect private intelligence and constrain authority. The central claim is architectural and falsifiable: durable autonomous work requires proof-carrying loops that prevent persuasive but unsupported output from becoming institutional memory. This canonical edition specifies the object model, security boundary, user experience, production artifacts, conformance levels, evaluation program, and first external proof run required to test that claim.
\end{abstract}

\begin{goldbox}[Executive Compression]
\textbf{The problem:} output is abundant; justified authority is scarce. \textbf{The mechanism:} Proof Debt becomes AGI Jobs; Evidence becomes Chronicle candidates; admitted skills become reusable capability. \textbf{The cryptographic boundary:} Merkle roots and ZK receipts commit proof state without exposing plaintext private intelligence. \textbf{The empirical burden:} one externally reviewed mission, one admitted skill, one verifiable graph root, and one measured improvement on a second mission.
\end{goldbox}

\textbf{Keywords:} GoalOS; proof loops; autonomous agents; Proof Mission; Proof Debt; AGI Jobs; ProofBundles; Evidence Docket; Chronicle; Validated Skill Graph; Merkle roots; zero-knowledge proofs; ENS; governed capability improvement; proof economy.

\tableofcontents
\clearpage

\input{modules/01_executive.tex}
\input{modules/02_doctrine_loop.tex}
\input{modules/03_proof_work.tex}
\input{modules/04_evidence_memory.tex}
\input{modules/05_validated_skill_graph.tex}
\input{modules/06_merkle_privacy_authority.tex}
\input{modules/07_governed_improvement.tex}
\input{modules/08_production_user_experience.tex}
\input{modules/09_evaluation_proof_run.tex}
\input{modules/10_appendices.tex}

\end{document}
